Cloud correction of humidity radiances with case-specific uncertainty estimate

### 2020-11-23 Article/ici_submm_cloudcorr.tex
fixed type-setting errors, suggestions from Vasilis

\newcommand{\todo}[1]{{\color{red} #1}}
\newcommand{\ynfcs}{y_\text{NFCS}}
\newcommand{\yonfcs}{y^{o}_\text{NFCS}}
\newcommand{\ynfcsi}{y_\text{NFCS, i}}
\newcommand{\yonfcsi}{y^{o}_\text{NFCS, i}}
\newcommand{\y}{\vec{y}}

@@

\affil[1]{Department of Space, Earth and Environment, Chalmers University of
  Technology, Gothenburg, Sweden} 
\affil[2]{European Centre for Medium Range Weather Forecasts, Reading, UK}

@@

A methodology based on quantile regression neural networks (QRNN) is presented that identifies and corrects the cloud impact on microwave humidity sounder radiances at 183\,GHz. This approach estimates the posterior distributions of noise free clear-sky (NFCS) radiances, providing nearly bias-free estimates of clear-sky radiances with a full posterior error distribution. It is first demonstrated by application to a present sensor, the MicroWave Humidity Sounder-2 (MWHS-2), then the applicability to sub-millimeter (sub-mm) sensors is also analysed. The QRNN results improve upon what operational cloud filtering techniques like a scattering index can achieve, but are ultimately imperfect due to limited information content on cirrus impact from traditional microwave channels---the negative departures associated with high cloud impact are successfully corrected, but thin cirrus clouds cannot be fully corrected. In contrast, when sub-mm observations are used, QRNN successfully corrects most cases with cloud impact, with only 2-6\% of cases left partially corrected. The methodology works well even if only one sub-mm channel (325\,GHz) is available. When using sub-mm observations, cloud correction usually results in error distributions with standard deviation less than typical channel noise values. Furthermore, QRNN outputs predicted quantiles for case-specific uncertainty estimates, successfully representing the uncertainty of cloud correction for each observation individually. In comparison to deterministic correction or filtering approaches, the corrected radiances and attendant uncertainty estimates have great potential to be used efficiently in assimilation systems due to being largely unbiased and adding little further uncertainty to the measurements.

@@

\introduction
%
Satellite observations of humidity inside the troposphere are mainly performed
by downward-looking sensors. Among this class of observations, the frequency
range around 183\,GHz has a special position. Water vapour has a noticeable
transition at 22\,GHz, but it is relatively weak and only column values can be
derived \citep[e.g.][]{schluessel1990atmospheric} for the observation geometry
of concern. The first transition in the microwave region that can be used to
derive altitude information, i.e.\ can be used for ``sounding'', is the one at
183.31\,GHz \citep{kakar1983retrieval,wang1983profiling}. On the other hand, at
infrared wavelengths a high number of water vapour transitions are found,
including some of high strength. As a consequence, infrared sounders can
provide humidity profiles with high precision and good vertical resolution, but
with strong limitations imposed by clouds. To be able to also sense humidity
inside and below clouds, weather satellites are since some time equipped with
channels around 183\,GHz. Today such channels are part of several sensors, such
as ATMS (Advanced Technology Microwave Sounder, \citet{weng2012introduction}).

Although microwave channels are less affected by cloud contamination, precipitation and most dense clouds, particularly if found at high altitude, can
still affect measured radiances around 183\,GHz
\citep[e.g.][]{bennartz2003sensitivity}. As the impact from the hydrometeors
then is dominated by scattering, the complexity of the analysis of the data
increases dramatically and there exists a need to identify the problematic
cases. This is normally denoted as cloud filtering, to obtain data of ``clear
sky'' character. Such filtering has been applied to derive climate records
\citep{lang2020new} and is essential in studies of the agreement between
observations and simulations \citep{brogniez2016review} as well as 
comparing observations of different instruments to validate their calibration
\citep{john2013assessment,moradi:retri:15,berg2016intercalibration}. A commonly
used cloud filtering methods for these applications is the one of
\citet{buehler:aclou:07}. This method is based on the 183\,GHz data alone,
involving rules on the brightness temperatures differences between channels. An
older version is \citet{burns1997effects}.

Another motivation necessitating the need for cloud clearing is usage of 183\,GHz channels in 
numerical weather prediction (NWP). Usage of passive microwave data by
``all-sky'' assimilation in global NWP is growing \citep{geer2017growing},
but 183\,GHz data are still mainly used in a clear-sky fashion
\citep{geer2018all}. The latter is particularly true in NWP of regional scope
\citep{gustafsson2018survey}, with clear-sky assimilation of 183GHz radiances still commonplace. Regardless, both clear-sky and all-sky assimilation require identification of cloud affected observations, either to screen out these observations or to assign an appropriate observation error. ``Scattering index'' \citep{geer2015scatteringindex} is one of the approaches used. It is based on brightness temperature differences between 89 and 150\,GHz. Another commonly used cloud filtering method is based on ``observation minus background'' (O-B), where the forecast model is used to obtain an estimate of the expected clear-sky value and the observation is rejected if the deviation exceeds some threshold \citep{English1999clouddetection}.

%Despite their broad usage, these filtering methods aim at partial removal of cloud contamination originating from high ice clouds. The residual cloud impact can have some drawbacks. 
At 183\,GHz, the impact of hydrometeors typically causes a decrease in the observed radiance due to scattering from ice hydrometeors (e.g.\ \citet{barlakas:three:20}). This implies that if any cloud contamination is missed by the filtering, a negative bias in mean radiance will result compared to the true clear-sky mean, which may translate into a bias in humidity after the retrieval or assimilation. For NWP systems assimilating clear-sky observations, the effect of undetected clouds may be overcome by inflating the observational errors and diminishing the impact of observations. Furthermore, the mathematical assumptions of data assimilation (DA) are predicated on Gaussian errors with no mean bias, and residual cloud impacts that cause a net bias are not easily handled by variational bias correction. One solution is to apply a very strict filtering, but this increases the rejection of clear-sky values, i.e.\ an important loss of useful data. Another limitation of existing clearing approaches is their ``one for all'' approach, i.e.\ observations in all 183\,GHz channels are either kept or rejected. This often rejects more observations than needed, as the channels differ in their altitude coverage. An observation could be cloudy in some channels and still the be clear-sky in others. To allow a channel specific filtering, data likely need to be combined in a more complex manner than simple differences, but it is unclear what type of regression would be best as the ideal solution would be scene-dependent. This points towards applying machine learning, as used by e.g.\ \citet{favrichon2019detecting}. A maybe less obvious problem is the assignment of uncertainty to the filtered values. To our best knowledge, so far only estimates of mean and worst case errors exist in the literature. Some cases with relatively high cloud impact will likely be missed, while most cases are clear-sky from the start. As the remaining cloudy cases can cause significant biases, the likely solution is to apply a quite conservative (high) error estimate. However, this will unnecessarily downgrade the value of the truly clear-sky cases and the observations are used in a non-optimal manner.

@@

The main focus of this study is the potential of this cloud-correction
method using sub-millimetre (sub-mm) observations, which will become available
operationally with the launch of the Ice Cloud Imager (ICI,
\citet{eriksson:towar:20}) on board the next generation of EUMETSAT Polar System - Second Generation (EPS-SG), but will likely also be included in
several smaller missions (such as Artic Weather Satellite (AWS); presented
below). Additionally, we demonstrate the feasibility of the approach based on
observations at 89 and 150\ GHz channels (following
\citet{geer2015scatteringindex}), which are available on several sensors extant today. The focus
on sub-mm channels is motivated by several reasons. First, the higher
frequencies are more sensitive to scattering effects from smaller
hydrometeors and are thus expected to provide greater sensitivity to high altitude cirrus clouds. For example, in some cloudy situations the
cloud impact at 183\,GHz may be of the order of thermal noise and modelling uncertainties,
while the impact at 325\,GHz is significant enough to provide sufficient signal to noise for identifying cloud. Second, the proposed
cloud correction methods allow integration of ICI sub-mm observations in
clear-sky DA schemes with no further modifications, thus providing a simple
way to make use this novel data source as soon as it becomes available.

@@

\section{Data and methods}
\label{data_methods}
%
\subsection{Satellite Instruments}
%
\subsubsection{ MicroWave Humidity Sounder-2}
%
%t
\begin{table}[t]
	\caption{Specifications of MWHS-2 channels relevant to this study.}
	\label{tab:specifications_MWHS2}	
	\begin{tabular}{crrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
		& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		1	&	89.0   		  & 1500			&	1.0	\\
		6	&	118.75±1.1    & \phantom{0}200 	&	1.6\\
		7	&	118.75±2.5    & \phantom{0}200 	&	1.6\\
		10	&	150.0         & 1500 			&	1.0 \\
		11	&	183.31±1.0      & \phantom{0}500  &	1.0 \\
		12  & 	183.31±1.8    & \phantom{0}700 	&   1.0\\
		13  & 	183.31±3.0      & 1000    		&	1.0	\\
		14  & 	183.31±4.5    & 2000    		&	1.0\\
		15  & 	183.31±7.0      & 2000  			&	1.0  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
The MicroWave Humidity Sounder 2 (MWHS-2) is an instrument on two current satellites in the FengYun-3 series, FY-3C and FY-3D. MWHS-2 a cross track scanning microwave radiometer and measures 15 frequencies in the range 89-191\,GHz. 89\,GHz and 150\,GHz are window channels, five humidity sounding channels are centered at 183.31\,GHz, and eight temperature sounding channels are centered on the 118.75\,GHz oxygen absorption line. The five humidity sounding channels are similar to ATMS. Observations from MWHS-2 are routinely assimilated in all-sky conditions at the European Centre for Medium-Range Weather Forecasts (ECMWF) with demonstrable positive impact on forecast performance \citep{duncan2020MWHS}. The channels relevant to this study are described in Table~\ref{tab:specifications_MWHS2}.

@@

The Ice Cloud Imager (ICI) is a new instrument on board EPS-SG satellite MetOp-SG (Meteorological Operational - Second Generation). MetOp-SG is scheduled for launch in 2024, and it will make ICI the first operational sensor observing Earth using sub-mm wavelengths. The main objective of ICI is to use high frequency channels for measuring ice cloud properties, and improve the representation of ice clouds in regional and global NWP models. ICI is a conically scanning radiometer that will measure 13 frequencies from 183\,GHz up to 664\, GHz.  Among all available channels, 183.31\,GHz, 325.15\,GHz and 448.00\,GHz, will measure vertical polarization;  while other channels around frequencies 243.20\,GHz and 664.00\,GHz are ``window channels'' and will measure both vertical and horizontal polarization. The instrument will observe earth from a mean altitude of 832\,km with sensor viewing angle 44.767$^\unit{^{\degree}}$(measured from nadir). For all the channels, the mean footprint size is about 15\,km, but the exact geo-location of samples differs. Therefore, a simultaneous utilization of data from different channels shall require remapping to a common footprint \citep{eriksson:towar:20}.

For this study, we conducted the forward simulations of the channels around: 183.31\,GHz, 325.15\,GHz, 448.00\,GHz and 664.00\,GHz (Table~\ref{tab:ICI_MWI_channels}). For brevity, we assume that all simulations are mapped to a common footprint.

\subsubsection{Artic Weather Satellite}
%
\begin{table}[t]
	\caption{Specifications of AWS channels relevant to this study.}
	\label{tab:specifications_AWS}	
	\begin{tabular}{lrrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
				& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		AWS-32	&	176.31    & 2000	&	0.45	\\
		AWS-33	&	178.81    & 2000 	&	0.45\\
		AWS-34	&	180.31    & 1000 	&	0.64\\
		AWS-35	&	181.51    & 1000 	&	0.64 \\
		AWS-36	&	182.31    & \phantom{0}500  &	0.88 \\
		AWS-41  & 325.15$\pm$6.60    & 2800 	 &0.60\\
		AWS-42  & 325.15$\pm$4.10    & 1800    &0.75	\\
		AWS-43  & 325.15$\pm$2.40    & 1200    &0.92\\
		AWS-44  & 325.15$\pm$1.20    & \phantom{0}800  &1.12  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
The Arctic Weather Satellite (AWS) is a small satellite mission approved as European Space Agency (ESA)
Earth Watch Programme Element. It is a small platform carrying a single across-track scanning microwave radiometer. It is planned as a small but cost-effective platform, which can supplement the information from other polar orbiting satellites. The first prototype mission is expected to be delivered around 2023. The exact channels onboard AWS are not yet set, but five 183\,GHz and four 325\,GHz channels are presently planned. A brief summary of these channels is provided in Table~\ref{tab:specifications_AWS}.

In this study, forward simulations of the channels around 183.31\,GHz and 325.15\,GHz were conducted.

\subsection{Simulations}
%
MWHS-2 simulated radiances during the period June-July 2020, are sourced from ECMWF. In the current version of the ECMWF Integrated Forecasting System (IFS), cycle 47R1 \citep{IFS47R1chap1}, clear-sky and all-sky radiative transfer are performed simultaneously for monitoring purposes, despite all humidity sounders being assimilated via all-sky exclusively. These side by side radiative transfer calculations on a large variety of model scenes from the IFS provides an ideal dataset for comparing radiances with and without cloud effects. Out of all the available observations during the period, we use data for latitudinal range $60^{\degree}$\,S to $60^{\degree}$\,N and satellite zenith angle less than $7.5^{\degree}$. With this filter we have approximately 290\,000 cases.

ICI and AWS frequencies are simulated with Atmospheric Radiative Transfer Simulator (ARTS, \citet{buehler:artst:18}). CloudSat \citep{Stephens2002cloudsat} profiles during August 2015 are randomly selected for forward simulations. The input data are restricted between $60\unit{^{\degree}S}$ to $60\unit{^{\degree}N}$, and surface is below 500\,m. Both clear-sky and all-sky scenarios  are simulated. The complete simulation setup is described in Appendix \ref{appendix:ARTS_setup}. For ICI and AWS, 220\,000 and 143\,000 cases are simulated, respectively. All AWS sensor viewing angles (from $0^{\degree}$ to $45^{\degree}$) are simulated, but the results described in this study are based on nadir viewing angle. Simulations for all three sensors are noise free, so to incorporate the measurement uncertainties, whenever needed, gaussian noise is added to according to the channel NE$\Delta$T (Table~\ref{tab:specifications_MWHS2} - Table~\ref{tab:specifications_AWS}).

@@

For this application, the predicted percentiles are chosen to be
$0.2\%, 3\%, 16\%, 50\%, 85\%, 97\%$ and $99.8\%$. For a Gaussian
distribution with mean $\mu$ and standard deviation $\sigma$, these quantiles
approximately correspond to $\mu -3\sigma, \mu-2\sigma, \mu-\sigma
, \mu, \mu + \sigma, \mu + 2\sigma, \mu + 3\sigma$ and thus allows
estimation of the $\pm 1, \pm 2$ and $\pm 3\sigma$ confidence intervals.

QRNN's are trained to minimize the mean of the sum of the quantile loss functions
%
\begin{align}
\mathcal{L}_\tau(y_\tau, y) &=
\begin{cases}
\tau|y - y_\tau|, & y_\tau < y \\ (1 - \tau)|y_\tau - y|, & \text{otherwise}
\\
\end{cases}
\end{align}
%
for all selected quantile fractions $\tau$, where $y_\tau$ is the predicted
quantile and $y$ the reference value from the training or test data. The
quantile loss is also used in this study as a performance criterion for the tuning of
the hyper-parameters of the QRNN (see Appendix~\ref{appendix:hyperparamter}). In
addition to the quantile loss, also the Continuously Ranked Probability Score
(CRPS) is considered. Given a predicted CDF
$F$ and the reference value $y$, the CRPS is defined as
%
\begin{align}
\text{CRPS}(F, y) = \int_{-\infty}^{\infty} \left (F(y') - y\right )^2\: dy'.
\end{align}
%
To compute the CRPS for a prediction from a QRNN, the predicted quantiles are
used to derive a piece-wise linear approximation of the CDF of the predicted
distribution. Note that CRPS is only used to evaluate hyper-parameter tuning.

The implementation of QRNN is similar to the one described in
\citet{pfreundschuh:aneur:18}, except that this version uses PyTorch
\citep{paszke2017automatic} instead of Keras \citep{chollet2015keras} to
implement the underlying neural network. The implementation is available as a
part of version of the Typhon software package \citep{typhonv08}. The major
challenge for implementing QRNN for the current application was to select high
performing neural network architecture. This was obtained through grid search
over different hyper-parameter configurations. The details are described in
Appendix~\ref{appendix:hyperparamter}

@@

	\item In the first experiment, we examine the performance of QRNN cloud correction with MWHS-2 window channels: 89 and 150\,GHz. In ECMWF NWP system, the differences between the observations of these two window channels are used to identify the cloud affected data for humidity sounding channels \citep{geer2015scatteringindex}. To investigate the potential impact of these two window channels in QRNN based cloud correction, the configuration QRNN-single is applied. In this experiment, the training inputs include all-sky brightness temperatures from target 183\,GHz channel, 89\,GHz and 150\,GHz. Both 89 and 150\,GHz are window channels so the land sea mask is also included as a training input. For example, for channel 14, the training inputs are all-sky brightness temperatures from channels 14, 1, 10 and land/sea mask. This combination is referred as 89+150\,GHz in the text.

	\item In the second experiment, we explore if few of low peaking channels of 118\,GHz could have any potential advantage in cloud correction. To explore their impact, QRNN-single is trained with data from target 183\,GHz, 89\,GHz, 150\,GHz, 118.75$\pm$1.1\,GHz (channel 6), 118.75$\pm$2.5\,GHz (channel 7) and land/sea mask. This combination is denoted as 89+150+118\,GHz

>>>>

\newcommand{\todo}[1]{{\color{red} #1}}
\newcommand{\ynfcs}{y_\text{NFCS}}
\newcommand{\yonfcs}{y^{\circ}_\text{NFCS}}
\newcommand{\ynfcsi}{y_\text{NFCS, i}}
\newcommand{\yonfcsi}{y^{\circ}_\text{NFCS, i}}
\newcommand{\y}{\vec{y}}

@@

\affil[1]{Department of Space, Earth and Environment, Chalmers University of
  Technology, Gothenburg, Sweden} 
\affil[2]{European Centre for Medium Range Weather Forecasts, Reading, United Kingdom}

@@

A methodology based on quantile regression neural networks (QRNN) is presented that identifies and corrects the cloud impact on microwave humidity sounder radiances at 183\,GHz. This approach estimates the posterior distributions of noise free clear-sky (NFCS) radiances, providing nearly bias-free estimates of clear-sky radiances with a full posterior error distribution. It is first demonstrated by application to a present sensor, the MicroWave Humidity Sounder-2 (MWHS-2), then the applicability to sub-millimeter (sub-mm) sensors is also analysed. The QRNN results improve upon what operational cloud filtering techniques like a scattering index can achieve, but are ultimately imperfect due to limited information content on cirrus impact from traditional microwave channels---the negative departures associated with high cloud impact are successfully corrected, but thin cirrus clouds cannot be fully corrected. In contrast, when sub-mm observations are used, QRNN successfully corrects most cases with cloud impact, with only 2--6\% of the cases left partially corrected. The methodology works well even if only one sub-mm channel (325\,GHz) is available. When using sub-mm observations, cloud correction usually results in error distributions with standard deviation less than typical channel noise values. Furthermore, QRNN outputs predicted quantiles for case-specific uncertainty estimates, successfully representing the uncertainty of cloud correction for each observation individually. In comparison to deterministic correction or filtering approaches, the corrected radiances and attendant uncertainty estimates have great potential to be used efficiently in assimilation systems due to being largely unbiased and adding little further uncertainty to the measurements.

@@

\introduction
%
Satellite observations of humidity inside the troposphere are mainly performed
by downward-looking sensors. Among this class of observations, the frequency
range around 183\,GHz has a special position. Water vapour has a noticeable
transition at 22\,GHz, but it is relatively weak and only column values can be
derived \citep[e.g.,][]{schluessel1990atmospheric} for the observation geometry
of concern. The first transition in the microwave region that can be used to
derive altitude information, i.e.\ ``sounding'', is the one at
183\,GHz \citep{kakar1983retrieval,wang1983profiling}. On the other hand, at
infrared wavelengths a high number of water vapour transitions are found,
including some of high strength. As a consequence, infrared sounders can
provide humidity profiles with high precision and good vertical resolution, but
with strong limitations imposed by clouds. To be able to also sense humidity
inside and below clouds, weather satellites are since some time equipped with
channels around 183\,GHz. Today such channels are part of several sensors, such
as ATMS \citep[Advanced Technology Microwave Sounder,][]{weng2012introduction}.

Although microwave channels are less affected by cloud contamination, precipitation and most dense clouds, particularly if found at high altitude, can
still affect measured radiances around 183\,GHz
\citep[e.g.][]{bennartz2003sensitivity}. As the impact from the hydrometeors
then is dominated by scattering, the complexity of the analysis of the data
increases dramatically and there exists a need to identify the problematic
cases. This is normally denoted as cloud filtering, to obtain data of ``clear
sky'' character. Such filtering has been applied to derive climate records
\citep{lang2020new} and is essential in studies of the agreement between
observations and simulations \citep{brogniez2016review} as well as 
comparing observations of different instruments to validate their calibration
\citep{john2013assessment,moradi:retri:15,berg2016intercalibration}. Commonly
used cloud filtering methods for these applications are based on 183\,GHz data alone, involving rules on the brightness temperatures differences between channels \citep{burns1997effects, buehler:aclou:07}.

Another motivation necessitating the need for cloud clearing is usage of 183\,GHz channels in 
numerical weather prediction (NWP). Usage of passive microwave data by
all-sky assimilation in global NWP is growing \citep{geer2017growing},
but 183\,GHz data are still mainly used in a clear-sky fashion
\citep{geer2018all}. The latter is particularly true in NWP of regional scope
\citep{gustafsson2018survey}, with clear-sky assimilation of 183\,GHz radiances still commonplace. Regardless, both clear-sky and all-sky assimilation require identification of cloud affected observations, either to screen out these observations or to assign an appropriate observation error. The most commonly used cloud filtering techniques are the ``Scattering index'' \citep{geer2015scatteringindex} and the ``observation minus background'' (O$-$B). The first one is based on brightness temperature differences between 89 and 150\,GHz. In the second one, the forecast model is used to obtain an estimate of the expected clear-sky value and the observation is rejected if the deviation exceeds some threshold \citep{English1999clouddetection}.

At 183\,GHz, the impact of hydrometeors typically causes a decrease in the observed radiance due to scattering from ice hydrometeors \citep[e.g.,][]{barlakas:three:20}. This implies that if any cloud contamination is missed by the filtering, a negative bias in the mean radiance, compared to the true clear-sky mean, may translate into a bias in humidity after the retrieval or assimilation. For NWP systems assimilating clear-sky observations, the effect of undetected clouds may be overcome by inflating the observational errors and diminishing the impact of observations. Furthermore, the mathematical assumptions of data assimilation (DA) are predicated Gaussian errors with no mean bias, and residual cloud impacts that cause a net bias are not easily handled by variational bias correction. One solution is to apply a very strict filtering, but this increases the rejection of clear-sky values, i.e.\ an important loss of useful data. Another limitation of existing filtering approaches is their ``one for all'' approach, i.e.\ observations in all 183\,GHz channels are either kept or rejected. This often rejects more observations than needed, as the channels differ in their altitude coverage. An observation could be cloudy in some channels and still the be clear-sky in others. To allow a channel specific filtering, data likely need to be combined in a more complex manner than simple differences, but it is unclear what type of regression would be best as the ideal solution would be scene-dependent. This points towards applying machine learning techniques \citep[e.g.,][]{favrichon2019detecting}. A maybe less obvious problem is the assignment of uncertainty to the filtered values. To our best knowledge, so far only estimates of mean and worst case errors exist in the literature. Some cases with relatively high cloud impact will likely be missed, while most cases are clear-sky from the start. As the remaining cloudy cases can cause significant biases, the likely solution is to apply a quite conservative (high) error estimate. However, this will unnecessarily downgrade the value of the truly clear-sky cases and the observations are used in a non-optimal manner.

@@

The main focus of this study is the potential of this cloud-correction
method using sub-millimetre (sub-mm) observations, which will become available
operationally with the launch of the Ice Cloud Imager \citep[ICI,][]{eriksson:towar:20} on board the next generation of European Organisation for the Exploitation of Meteorological Satellites Polar System - Second Generation (EUMETSAT EPS-SG), but will likely also be included in
several smaller missions (such as Arctic Weather Satellite (AWS); presented
below). Additionally, we demonstrate the feasibility of the approach based on
89 and 150\ GHz channels \citep[following][]{geer2015scatteringindex}, which are available on several sensors extant today. The focus
on sub-mm channels is motivated by several reasons. First, the higher
frequencies are more sensitive to scattering effects from smaller
hydrometeors and are thus expected to provide greater sensitivity to high altitude cirrus clouds. For example, in some cloudy situations the
cloud impact at 183\,GHz may be of the order of thermal noise and modelling uncertainties,
while the impact at 325\,GHz is significant enough to provide sufficient signal to noise for identifying cloud. Second, the proposed
cloud correction methods allow integration of ICI sub-mm observations in
clear-sky DA schemes with no further modifications, thus providing a simple
way to make use this novel data source as soon as it becomes available.

@@

\section{Data and methods}
\label{data_methods}
%
\subsection{Satellite Instruments}
%
\subsubsection{ MicroWave Humidity Sounder-2}
%
%t
\begin{table}[t]
	\caption{Specifications of MWHS-2 channels relevant to this study.}
	\label{tab:specifications_MWHS2}	
	\begin{tabular}{crrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
		& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		1	&	89.0   		  & 1500			&	1.0	\\
		6	&	118.75±1.1    & \phantom{0}200 	&	1.6\\
		7	&	118.75±2.5    & \phantom{0}200 	&	1.6\\
		10	&	150.0         & 1500 			&	1.0 \\
		11	&	183.31±1.0      & \phantom{0}500  &	1.0 \\
		12  & 	183.31±1.8    & \phantom{0}700 	&   1.0\\
		13  & 	183.31±3.0      & 1000    		&	1.0	\\
		14  & 	183.31±4.5    & 2000    		&	1.0\\
		15  & 	183.31±7.0      & 2000  			&	1.0  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
The MicroWave Humidity Sounder 2 (MWHS-2) is an instrument on two current satellites in the FengYun-3 series, FY-3C and FY-3D. MWHS-2 a cross track scanning microwave radiometer and measures 15 frequencies in the range 89--191\,GHz. 89\,GHz and 150\,GHz are window channels, five humidity sounding channels are centered around 183\,GHz, and eight temperature sounding channels are centered on the 118.75\,GHz oxygen absorption line. The five humidity sounding channels are similar to ATMS. Observations from MWHS-2 are routinely assimilated in all-sky conditions at the European Centre for Medium-Range Weather Forecasts (ECMWF) with demonstrable positive impact on forecast performance \citep{duncan2020MWHS}. The channels relevant to this study are described in Table~\ref{tab:specifications_MWHS2}.

@@

The ICI is a new instrument on board EPS-SG satellite MetOp-SG (Meteorological Operational - Second Generation). MetOp-SG is scheduled for launch in 2024, and it will make ICI the first operational sensor observing Earth using sub-mm wavelengths. The main objective of ICI is to use high frequency channels for measuring ice cloud properties, and improve the representation of ice clouds in regional and global NWP models. ICI is a conically scanning radiometer that will measure 13 frequencies from 183\,GHz up to 664\,GHz.  Among all available channels, 183\,GHz, 325\,GHz and 448\,GHz, will measure vertical polarization;  while other channels around 243\,GHz and 664\,GHz are ``window channels'' and will measure both vertical and horizontal polarization. The instrument will observe Earth from a mean altitude of 832\,km with the sensor viewing angle 44.767$^\unit{^{\degree}}$(measured from nadir). For all the channels, the mean footprint size is about 15\,km, but the exact geo-location of samples differs. Therefore, a simultaneous utilization of data from different channels shall require remapping to a common footprint \citep{eriksson:towar:20}.

For this study, we conducted the forward simulations of the channels around: 183\,GHz, 325\,GHz, 448\,GHz and 664\,GHz (Table~\ref{tab:ICI_MWI_channels}). For brevity, we assume that all simulations are mapped to a common footprint.

\subsubsection{Artic Weather Satellite}
%
\begin{table}[t]
	\caption{Specifications of AWS channels relevant to this study.}
	\label{tab:specifications_AWS}	
	\begin{tabular}{lrrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
				& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		AWS-32	&	176.31    & 2000	&	0.45	\\
		AWS-33	&	178.81    & 2000 	&	0.45\\
		AWS-34	&	180.31    & 1000 	&	0.64\\
		AWS-35	&	181.51    & 1000 	&	0.64 \\
		AWS-36	&	182.31    & \phantom{0}500  &	0.88 \\
		AWS-41  & 325.15$\pm$6.60    & 2800 	 &0.60\\
		AWS-42  & 325.15$\pm$4.10    & 1800    &0.75	\\
		AWS-43  & 325.15$\pm$2.40    & 1200    &0.92\\
		AWS-44  & 325.15$\pm$1.20    & \phantom{0}800  &1.12  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
The AWS is a small satellite mission approved as European Space Agency (ESA) Earth Watch Programme Element. It is a small platform carrying a single across-track scanning microwave radiometer. It is planned as a small but cost-effective platform, which can supplement the information from other polar orbiting satellites. The first prototype mission is expected to be delivered around 2023. The exact channels onboard AWS are not yet set, but five 183\,GHz and four 325\,GHz channels are presently planned. A brief summary of these channels is provided in Table~\ref{tab:specifications_AWS}.

In this study, forward simulations of the channels around 183\,GHz and 325\,GHz were conducted.

\subsection{Simulations}
%
MWHS-2 simulated radiances during the period June--July 2020 are sourced from ECMWF. In the current version of the ECMWF Integrated Forecasting System (IFS), cycle 47R1 \citep{IFS47R1chap1}, clear-sky and all-sky radiative transfer are performed simultaneously for monitoring purposes, despite all humidity sounders being assimilated via all-sky exclusively. These side by side radiative transfer calculations on a large variety of model scenes provides an ideal dataset for comparing radiances with and without cloud effects. Out of all the available observations during the period, we use data for the latitudinal range: $60^{\degree}$\,S to $60^{\degree}$\,N and satellite zenith angle, less than $7.5^{\degree}$. With this filter, we have approximately 290\,000 cases.

ICI and AWS frequencies are simulated with Atmospheric Radiative Transfer Simulator \citep[ARTS,][]{buehler:artst:18}. For the forward simulations, Cloud Satellite \citep[Cloudsat,][]{Stephens2002cloudsat} profiles during August 2015 are randomly selected. The input data are restricted between $60\unit{^{\degree}S}$ to $60\unit{^{\degree}N}$, and surface is below 500\,m. Both clear-sky and all-sky scenarios  are simulated. The complete simulation setup is described in Appendix \ref{appendix:ARTS_setup}. For ICI and AWS, 220\,000 and 143\,000 cases are simulated, respectively. All AWS sensor viewing angles (from $0^{\degree}$ to $45^{\degree}$) are simulated, but the results described in this study are based on nadir viewing angle. Simulations for all three sensors are noise free, so to incorporate the measurement uncertainties, whenever needed, Gaussian noise is added according to the channel NE$\Delta$T (Table~\ref{tab:specifications_MWHS2} -- Table~\ref{tab:specifications_AWS}).

@@

For this application, the predicted percentiles are chosen to be
$0.2\,\%, 3\,\%, 16\,\%, 50\,\%, 85\,\%, 97\,\%$ and $99.8\,\%$. For a Gaussian
distribution with mean $\mu$ and standard deviation $\sigma$, these quantiles
approximately correspond to $\mu -3\sigma, \mu-2\sigma, \mu-\sigma
, \mu, \mu + \sigma, \mu + 2\sigma, \mu + 3\sigma$ and thus allows
estimation of the $\pm\, 1, \pm\, 2$ and $\pm\, 3\sigma$ confidence intervals.

QRNN's are trained to minimize the mean of the sum of the quantile loss functions,
%
\begin{align}
\mathcal{L}_\tau(y_\tau, y) &=
\begin{cases}
\tau|y - y_\tau|, & y_\tau < y \\ (1 - \tau)|y_\tau - y|, & \text{otherwise}
\\
\end{cases}
\end{align}
%
for all selected quantile fractions $\tau$, where $y_\tau$ is the predicted
quantile and $y$ the reference value from the training or test data. The
quantile loss is also used in this study as a performance criterion for the tuning of
the hyper-parameters of the QRNN (see Appendix~\ref{appendix:hyperparamter}). In
addition to the quantile loss, also the Continuously Ranked Probability Score
(CRPS) is considered. Given a predicted CDF
$F$ and the reference value $y$, the CRPS is defined as
%
\begin{align}
\text{CRPS}(F, y) = \int_{-\infty}^{\infty} \left (F(y') - y\right )^2\: dy'.
\end{align}
%
To compute the CRPS for a prediction from a QRNN, the predicted quantiles are
used to derive a piece-wise linear approximation of the CDF of the predicted
distribution. Note that CRPS is only used to evaluate hyper-parameter tuning.

The implementation of QRNN is similar to the one described in
\citet{pfreundschuh:aneur:18}, except that this version uses PyTorch
\citep{paszke2017automatic} instead of Keras \citep{chollet2015keras} to
implement the underlying neural network. The implementation is available as a
part of version of the Typhon software package \citep{typhonv08}. The major
challenge for implementing QRNN for the current application was to select high
performing neural network architecture. This was obtained through grid search
over different hyper-parameter configurations. The details are described in
Appendix~\ref{appendix:hyperparamter}.

@@

	\item In the first experiment, we examine the performance of QRNN cloud correction with MWHS-2 window channels: 89 and 150\,GHz. In ECMWF NWP system, the differences between the observations of these two window channels are used to identify the cloud affected data for humidity sounding channels \citep{geer2015scatteringindex}. To investigate the potential impact of these two window channels in QRNN based cloud correction, the configuration QRNN-single is applied. The training inputs include all-sky brightness temperatures from target 183\,GHz channel, 89\,GHz and 150\,GHz. Both 89 and 150\,GHz are window channels so the land sea mask is also included as a training input. For example, for channel 14, the training inputs are all-sky brightness temperatures from channels 14, 1, 10 and land/sea mask. This combination is referred as 89+150\,GHz in the text.

	\item In the second experiment, we explore if few of low peaking channels of 118\,GHz could have any potential in cloud correction. To explore their impact, QRNN-single is trained with data from target 183\,GHz, 89\,GHz, 150\,GHz, 118.75$\pm$1.1\,GHz (channel 6), 118.75$\pm$2.5\,GHz (channel 7) and land/sea mask. This combination is denoted as 89+150+118\,GHz.
